Supporting Information for Coherent IR-Hyper-Raman Four Wave Mixing Spectroscopy

### 2024-09-20 si.tex
revise SI

\section{Evaluating Herzberg-Teller Integrals for Harmonic Wells of Identical Curvature}
In this section, we discuss the evaluation of Herzberg-Teller integrals, in the limit of no Duschinsky coupling, when the ground and first excited state are harmonic wells of identical curvature.\cite{HerzbergTeller1933, Duschinsky1937}
We investigate the one dimensional case using the Hamiltonian described in the main text, $H = H_g |g) \left(g| + H_e |e\right) (e|$, where
\begin{subequations}\label{Hamiltonian}
	\begin{equation}
		H_g = \frac{\hbar \omega }{2} \left(p^2 + q^2 \right)
	\end{equation}
	and
	\begin{equation}
		H_e = \frac{\hbar \omega }{2} \left(p^2 +  (q-\Delta)^2 \right) + \hbar \omega_{eg}.
	\end{equation} 
\end{subequations}
Here, $p,q$ are the dimensionless conjugate normal mode coordinates, $\hbar\omega_{eg}$ is the energy difference between $H_e$ and $H_g$, and $\Delta$ is the dimensionless offset of the excited state surface relative to the ground state.
Franck-Condon factors and Herzberg-Teller integrals are often evaluated in position space using known solutions to the Harmonic Oscillator Hamiltonian, 
\begin{equation}
	\psi_m(q) = \langle q | m \rangle = N(m) H_m(q) \exp(-q^2/2),
\end{equation}
where $N(m) = (m! \ 2^m \pi^{\frac{1}{2}})^{-\frac{1}{2}}$ is the normalization factor and $H_m$ is the m$^{th}$ Hermite Polynomial.\cite{RN230}
We begin by generating an expression for $\mel{\tilde{n}}{q^k}{m}$, as this will resolve into Franck-Condon factors for $k = 0$ and Herzberg-Teller integrals for $k=1$.
The relevant Franck-Condon factors and Herzberg-Teller integrals, assuming identical curvature and no Duschinsky coupling, are plotted in \autoref{fig:fcht} as a function of offset, $\Delta$.
As in the main text, $\{\ket{\tilde{a}}\}$ are vibrational states on the first excited state $|e)$ and $\{\ket{b}\}$ are vibrational states on the ground state $|g)$.

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=6.675in]{figures/fcht.png}
	\caption{Plots of select (a) Franck-Condon factors and (b) Herzberg-Teller integrals relevant to this manuscript as a function of offset, $\Delta$.} 
	\label{fig:fcht}
\end{figure}

By letting $\ket{m}$ belong to the ground state surface $|g)$ and $\ket{\tilde{n}}$ belong to the first excited surface $|e)$, we write
\begin{equation}
\begin{split}
		\mel{\tilde{n}}{q^k}{m} &= \int_{-\infty}^\infty \mathrm{d}q \psi^*_n(q-\Delta) q^k \psi_m(q)\\
		&= \int_{-\infty}^\infty \mathrm{d}q N(n) N(m) H_n(q-\Delta) \exp(-\frac{(q - \Delta)^2}{2}) q^k H_m(q) \exp(-\frac{q^2}{2}),
\end{split}
\end{equation}
where $k=0$ corresponds to a Franck-Condon factor and $k=1$ refers to a Herzberg-Teller integral.
Introducing $\tilde{N} \equiv N(n)N(m)$ and slightly rearranging, we see
\begin{equation}\label{key}
	\mel{\tilde{n}}{q^k}{m} = \tilde{N} \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q H_n(q-\Delta) q^k H_m(q) \exp(-q^2 + q\Delta)
\end{equation}
This is the key result.
While integrals of form \autoref{key} can be computed straightforwardly using mathematical packages such as \texttt{Mathematica} or \texttt{SymPy}, \cite{Mathematica, sympy} it is a useful exercise in Gaussian integral manipulation to evaluate Herzberg-Teller or Franck-Condon integrals.
As a side note, \autoref{key} explicitly shows
\begin{subequations}
	\begin{equation}
		\lim_{\Delta \rightarrow 0} \langle \tilde{n} | m \rangle = \langle n | m \rangle = \delta_{n,m}
	\end{equation}
and
	\begin{equation}
	\lim_{\Delta \rightarrow 0} \mel{\tilde{n}}{q}{m} = \mel{{n}}{q}{m} = \sqrt{\frac{1}{2}} \left(\sqrt{m}\delta_{n,m-1} + \sqrt{m+1} \delta_{n,m+1}\right)
	\end{equation}
\end{subequations}
i.e., as the offset vanishes, the normal harmonic oscillator selection rules are obtained, the expected result, where $\delta$ is the Kronecher delta.

It is well known that the Hermite polynomials $H_v(q)$ are linear combinations of $q^w$, $w \in \mathbf{N}$. 
Thus, evaluating \autoref{key} for arbitrary $m,n,k$ values resolves into evaluating sums of integrals of the form
\begin{equation}\label{integral}
	\int_{-\infty}^\infty \mathrm{d}q \ q^w \exp(-q^2 + q\Delta)
\end{equation}
When $w=0$, the result follows by completing the square:\footnote{$\int_{-\infty}^{\infty} \mathrm{d}x \exp(-(x-a)^2) = \sqrt{\pi}$}
\begin{equation}
	\begin{split}
		\int_{-\infty}^\infty \mathrm{d}q \exp(-q^2 + q\Delta) &= \int_{-\infty}^\infty \mathrm{d}q \exp(-(q^2 - q\Delta +\frac{\Delta^2}{4} - \frac{\Delta^2}{4}))\\
		&= \exp(\frac{\Delta^2}{4}) \int_{-\infty}^\infty \mathrm{d}q \exp(-(q - \frac{\Delta}{2})^2) \\
		&= \sqrt{\pi} \exp(\frac{\Delta^2}{4})
	\end{split}
\end{equation}
 For $w \neq 0$, \autoref{integral} is evaluated through repeated use of a well-known Gaussian identity
\begin{equation}\label{identity}
	\int_{-\infty}^\infty \mathrm{d}q \ q^w \exp(-q^2 + q\Delta) = \partial_{\Delta} \int_{-\infty}^\infty \mathrm{d}q \ q^{w-1} \exp(-q^2 + q\Delta)
\end{equation}

We now have the background needed to calculate $\mel{\tilde{1}}{q}{0}$ for an arbitrary offset $\Delta$.
We note that Myers et al. have developed a general expression to evaluate Franck-Condon factors for two displaced harmonic wells of arbitrary curvature,\cite{Myers1982} so we focus our attention onto Herzberg-Teller integrals, which are not commonly tabulated.
Starting with \autoref{key},
\begin{equation}
\begin{split}
	\mel{\tilde{1}}{q}{0} = \tilde{N} \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q H_1(q-\Delta) q H_0(q) \exp(-q^2 + q\Delta)
\end{split}
\end{equation}
Using $H_0(q) = 1$ and $H_1(q) = 2q$,
\begin{equation}\label{HT10}
	\begin{split}
		\mel{\tilde{1}}{q}{0} &= \tilde{N} \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q \ 2(q - \Delta) q \exp(-q^2 + q\Delta)\\
		&= 2\tilde{N} \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q \ (q^2 - q \Delta) \exp(-q^2 + q\Delta)\\
		&=  2\tilde{N} \exp(-\frac{\Delta^2}{2}) \left(\int_{-\infty}^\infty \mathrm{d}q \ q^2 \exp(-q^2 + q\Delta) - \Delta \int_{-\infty}^\infty \mathrm{d}q \ q \exp(-q^2 + q\Delta) \right)
	\end{split}
\end{equation}
These integrals are now easily evaluated using \autoref{identity}:
\begin{subequations}\label{evalGaussians}
	\begin{equation}
		\begin{split}
					\int_{-\infty}^\infty \mathrm{d}q \ q \exp(-q^2 + q\Delta) &= \partial_{\Delta} \left(\sqrt{\pi} \exp(\frac{\Delta^2}{4})\right)\\
					& = \sqrt{\pi} \frac{\Delta}{2} \exp(\frac{\Delta^2}{4})
		\end{split}
	\end{equation}
giving
\begin{equation}
	\begin{split}
		\int_{-\infty}^\infty \mathrm{d}q \ q^2 \exp(-q^2 + q\Delta) &= \partial_{\Delta} \left(\sqrt{\pi} \left(\frac{\Delta}{2} \exp(\frac{\Delta^2}{4})\right)\right)\\
		& =\sqrt{\pi} \left(\frac{1}{2}\exp(\frac{\Delta^2}{4}) + \frac{\Delta^2}{4} \exp(\frac{\Delta^2}{4})\right)
	\end{split}
\end{equation}
\end{subequations}
With $\tilde{N} = N(1)*N(0) = (1! \ 2^1 \pi^{\frac{1}{2}})^{-\frac{1}{2}} * (0! \ 2^0 \pi^{\frac{1}{2}})^{-\frac{1}{2}} = \sqrt{1/2\pi}$ and substituting the results of \autoref{evalGaussians} into \autoref{HT10}, 
\begin{equation}
	\mel{\tilde{1}}{q}{0} = \sqrt{\frac{1}{2}} (1 - \frac{\Delta^2}{2}) \exp(\frac{-\Delta^2}{4})
\end{equation}
The evaluation of other Herzberg-Teller integrals follows a similar process.

@@

\section{Impact of $\Delta$ on $A,B$ Terms in HDFG Spectra}
In the main text, the potential well offset ($\Delta$) was taken as $\Delta = 0.5$.
Here, we vary $\Delta$ to investigate its impact on the structure of the $\abs{A+B}$ spectrum.

@@

\section{Evaluating impact of inhomogeneity on HDFG output}
In this section, the integral
\begin{equation}\label{generalcontour}
	\begin{split}
		\gamma_{ijkl} &= \int_{-\infty}^\infty \mathrm{d}\xi P(\xi) \gamma_{ijkl}(\xi)\\
		&= \frac{\sigma}{\pi}\int_{-\infty}^\infty \mathrm{d}\xi \frac{1}{\sigma^2 + \xi^2} \frac{\eta_{ijkl}}{\left(\Delta_{ga} - \xi\right)\left(\Delta_{ba}+ c\xi\right)} \\
		&= \frac{\eta_{ijkl} \sigma}{\pi} \int_{-\infty}^\infty \mathrm{d}\xi\frac{1}{(\xi + i\sigma)(\xi - i\sigma)} \frac{1}{\Delta_{ga} - \xi} \frac{1}{\Delta_{ba} + c\xi}\\
	\end{split}
\end{equation}
is evaluated using Contour integration.

@@

\subsection{Correlated Modes}
For correlated modes, the poles are situated in the complex plane as shown in \autoref{fig:contours}b.
It is clear that no matter which contour is chosen, there is a contribution from either $\Delta_{ba}$ or $\Delta_{ga}$. 
For the below calculation, we choose to evaluate along the contour in the bottom half of the complex plane, i.e., the poles are $\{-i\sigma, \Delta_{ga}\}$.
We stress that since the integral is evaluated over $\mathbb{R}$ and there are no degenerate poles, the results below can \textit{also} be obtained by the contour that wraps around the poles $\{i\sigma, -\Delta_{ba}\}$. 
Nevertheless, we choose the contour as drawn in \autoref{fig:contours}b.
Since the $\xi_k$ are nondegenerate, we see
\begin{widetext}
	\begin{equation}
		\begin{split}
			\int_{-\infty}^\infty \mathrm{d}\xi P(\xi) \gamma_{ijkl}(\xi) &= 2\pi i \sum_k \lim_{\xi \rightarrow \xi_k} P(\xi) \gamma_{ijkl}(\xi) (\xi - \xi_k)\\
			&= 2\pi i \frac{\eta_{ijkl} \sigma}{\pi}  \lim_{\xi \rightarrow -i\sigma} \frac{1}{(\xi + i\sigma)(\xi - i\sigma)} \frac{1}{\Delta_{ga} + \xi} \frac{1}{\Delta_{ba} - \xi} \left(\xi + i \sigma\right) \\ 
			&+ 2\pi i \frac{\eta_{ijkl} \sigma}{\pi} \lim_{\xi \rightarrow \Delta_{ga}} \frac{1}{(\xi + i\sigma)(\xi - i\sigma)} \frac{1}{\Delta_{ga} - \xi} \frac{1}{\Delta_{ba} - \xi} \left(\xi - \Delta_{ga}\right)\\
			&= 2 \eta_{ijkl} i \sigma \left(\frac{1}{2 i \sigma} \frac{1}{-i \sigma - \Delta_{ga}} \frac{1}{\Delta_{ba} - i\sigma} \right) - 2\eta_{ijkl} i \sigma \left(\frac{1}{\sigma^2 + \Delta_{ga} ^2} \frac{1}{\Delta_{ga} + \Delta_{ba}} \right)\\
			&= -\frac{\eta_{ijkl}}{\Delta_{ga} + i \sigma} \left(\frac{1}{\Delta_{ba} - i \sigma} + \frac{2i\sigma}{(\Delta_{ga} - i \sigma)(\Delta_{ga} + \Delta_{ba})}\right)\\
		\end{split}
	\end{equation}
\end{widetext}
Generally, when starting from the ground state, non-parametric spectroscopies line-narrow correlated modes, and parametric spectroscopies line-narrow anti-correlated modes. \cite{Dick83_1, RN425}

>>>>

\section{Evaluating Herzberg-Teller Integrals for Harmonic Wells of Identical Curvature}
In this section, we discuss the evaluation of Herzberg-Teller integrals when the ground and first excited state are harmonic wells of identical curvature.\cite{HerzbergTeller1933}
% We note that Myers et al. have developed a general expression to evaluate Franck-Condon factors for two displaced harmonic wells of differing curvature,\cite{Myers1982} 
We investigate the one dimensional case using the Hamiltonian described in the main text, $H = H_g |g) \left(g| + H_e |e\right) (e|$, where
\begin{subequations}\label{Hamiltonian}
	\begin{equation}
		H_g = \frac{\hbar \omega }{2} \left(p^2 + q^2 \right)
	\end{equation}
	and
	\begin{equation}
		H_e = \frac{\hbar \omega }{2} \left(p^2 +  (q-\Delta)^2 \right) + \hbar \omega_{eg}.
	\end{equation} 
\end{subequations}
Here, $p,q$ are the dimensionless conjugate normal mode coordinates, $\hbar\omega_{eg}$ is the energy difference between $H_e$ and $H_g$, and $\Delta$ is the dimensionless offset of the excited state surface relative to the ground state.
The Hamiltonians are the well-known quantum harmonic oscillator, and their solutions are
\begin{equation}
	\psi_m(q) = \langle q | m \rangle = N_m H_m(q) \exp(-q^2/2),
\end{equation}
where $N_m = (m! \ 2^m \pi^{\frac{1}{2}})^{-\frac{1}{2}}$ is the normalization factor and $H_m$ is the $m^{th}$ Hermite Polynomial.\cite{RN230}

For an $n$-quantum excited state vibration $| \tilde{n} \rangle$, and a $m$-quantum ground state vibration $| \tilde{m} \rangle$, the $k^{th}$-order vibrational coupling term is
\begin{equation}
	\begin{split}
			\mel{\tilde{n}}{q^k}{m} &= \int_{-\infty}^\infty \mathrm{d}q \ \psi^*_n(q-\Delta) q^k \psi_m(q)\\
			&= \int_{-\infty}^\infty \mathrm{d}q \ N_n H_n(q-\Delta) \exp(-\frac{(q - \Delta)^2}{2}) q^k N_m H_m(q) \exp(-\frac{q^2}{2}). \\
			&= N_n N_m \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q \ H_n(q-\Delta) q^k H_m(q) \exp(-q^2 + q\Delta)
		\end{split}\label{key}
\end{equation}
The integral corresponds to Franck-Condon matrix elements for $k=0$ and Herzberg-Teller matrix elements for $k=1$.

The matrix elements of \autoref{key} can be computed numerically, but analytic forms of the integral also exist. 
Due to the polynomial nature of $H_v(q)$, evaluation of \autoref{key} reduces into evaluating sums of integrals of the form
\begin{equation}\label{integral}
	\int_{-\infty}^\infty \mathrm{d}q \ q^w \exp(-q^2 + q\Delta)
\end{equation}
When $w=0$, the result follows by completing the square:\footnote{$\int_{-\infty}^{\infty} \mathrm{d}x \exp(-(x-a)^2) = \sqrt{\pi}$}
\begin{equation}
	\begin{split}
		\int_{-\infty}^\infty \mathrm{d}q \exp(-q^2 + q\Delta) &= \int_{-\infty}^\infty \mathrm{d}q \exp(-(q^2 - q\Delta +\frac{\Delta^2}{4} - \frac{\Delta^2}{4}))\\
		&= \exp(-\frac{\Delta^2}{4}) \int_{-\infty}^\infty \mathrm{d}q \exp(-(q - \frac{\Delta}{2})^2) \\
		&= \sqrt{\pi} \exp(-\frac{\Delta^2}{4})
	\end{split}
\end{equation}
 For $w \neq 0$, \autoref{integral} is evaluated through repeated use of a well-known Gaussian identity
\begin{equation}\label{identity}
	\int_{-\infty}^\infty \mathrm{d}q \ q^w \exp(-q^2 + q\Delta) = \partial_{\Delta} \int_{-\infty}^\infty \mathrm{d}q \ q^{w-1} \exp(-q^2 + q\Delta)
\end{equation}

We now have the background needed to calculate $\mel{\tilde{1}}{q}{0}$ for an arbitrary offset $\Delta$.
Using $H_0(q) = 1$ and $H_1(q) = 2q$,
\begin{equation}\label{HT10}
	\begin{split}
		\mel{\tilde{1}}{q}{0} &= N_1 N_0 \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q \ 2(q - \Delta) q \exp(-q^2 + q\Delta)\\
		&= 2N_1 N_0 \exp(-\frac{\Delta^2}{2}) \left( \int_{-\infty}^\infty \mathrm{d}q \ q^2 \exp(-q^2 + q\Delta) + \Delta \int_{-\infty}^\infty \mathrm{d}q \ q \exp(-q^2 + q\Delta) \right)
	\end{split}
\end{equation}
Both integrals can be evaluated using \autoref{identity}:
\begin{subequations}\label{evalGaussians}
	\begin{equation}
		\begin{split}
					\int_{-\infty}^\infty \mathrm{d}q \ q \exp(-q^2 + q\Delta) &= \partial_{\Delta} \left(\sqrt{\pi} \exp(\frac{\Delta^2}{4})\right)\\
					& = \sqrt{\pi} \frac{\Delta}{2} \exp(\frac{\Delta^2}{4})
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
		\int_{-\infty}^\infty \mathrm{d}q \ q^2 \exp(-q^2 + q\Delta) &= \partial_{\Delta} \left(\sqrt{\pi} \left(\frac{\Delta}{2} \exp(\frac{\Delta^2}{4})\right)\right)\\
		& =\sqrt{\pi} \left(\frac{1}{2}\exp(\frac{\Delta^2}{4}) + \frac{\Delta^2}{4} \exp(\frac{\Delta^2}{4})\right)
	\end{split}
\end{equation}
\end{subequations}
Substituting the results of \autoref{evalGaussians} into \autoref{HT10}, 
\begin{equation}
	\mel{\tilde{1}}{q}{0} = \sqrt{\frac{1}{2}} (1 - \frac{\Delta^2}{2}) \exp(\frac{-\Delta^2}{4})
\end{equation}
The evaluation of other Herzberg-Teller integrals is analogous.

\pagebreak
\section{Impact of displacement, resonance frequencies, and vibronic linewidth and on HDFG Spectra}
\begin{figure}[!htbp]
	\centering
	\includegraphics[width=6.675in]{figures/fcht.png}
	\caption{Plots of select (a) Franck-Condon factors and (b) Herzberg-Teller integrals relevant to this manuscript as a function of offset, $\Delta$.} 
	\label{fig:fcht}
\end{figure}

@@

\pagebreak

In the main text, the potential well offset ($\Delta$) was taken as $\Delta = 0.5$.
Here, we vary $\Delta$ to investigate its impact on the structure of the $\abs{A+B}$ spectrum.
The dependence of relevant Franck-Condon factors and Herzberg-Teller factors on displacement are plotted in \autoref{fig:fcht} as a function of offset, $\Delta$.
\autoref{fig:fcht} shows how different displacements, vibrational frequencies, and vibronic linewidths will alter the vibronic spectrum.

@@

\pagebreak

\section{Evaluating impact of inhomogeneity on HDFG output}
In this section, the integral
\begin{equation}\label{generalcontour}
	\begin{split}
		\gamma_{ijkl} &= \int_{-\infty}^\infty \mathrm{d}\xi P(\xi) \gamma_{ijkl}(\xi)\\
		&= \frac{\sigma}{\pi}\int_{-\infty}^\infty \mathrm{d}\xi \frac{1}{\sigma^2 + \xi^2} \frac{\eta_{ijkl}}{\left(\Delta_{ga} - \xi\right)\left(\Delta_{ba}+ c\xi\right)} \\
		&= \frac{\eta_{ijkl} \sigma}{\pi} \int_{-\infty}^\infty \mathrm{d}\xi\frac{1}{(\xi + i\sigma)(\xi - i\sigma)} \frac{1}{\Delta_{ga} - \xi} \frac{1}{\Delta_{ba} + c\xi}\\
	\end{split}
\end{equation}
is evaluated for $c=-1$ (electronic state and vibrational state energy fluctuations are anti-correlated) $c=1$ (electronic state and vibrational state energy fluctuations are correlated).

@@

\subsection{Correlated Modes}
For correlated modes, the poles are situated in the complex plane as shown in \autoref{fig:contours}b.
It is clear that no matter which contour is chosen, the contour will envelop two poles.
We choose to evaluate along the contour in the bottom half of the complex plane (see \autoref{fig:contours}b), where the poles are $\{-i\sigma, \Delta_{ga}\}$.
Since the $\xi_k$ are nondegenerate, we see
\begin{widetext}
	\begin{equation}
		\begin{split}
			\int_{-\infty}^\infty \mathrm{d}\xi P(\xi) \gamma_{ijkl}(\xi) &= 2\pi i \sum_k \lim_{\xi \rightarrow \xi_k} P(\xi) \gamma_{ijkl}(\xi) (\xi - \xi_k)\\
			&= 2\pi i \frac{\eta_{ijkl} \sigma}{\pi}  \lim_{\xi \rightarrow -i\sigma} \frac{1}{(\xi + i\sigma)(\xi - i\sigma)} \frac{1}{\Delta_{ga} + \xi} \frac{1}{\Delta_{ba} - \xi} \left(\xi + i \sigma\right) \\ 
			&+ 2\pi i \frac{\eta_{ijkl} \sigma}{\pi} \lim_{\xi \rightarrow \Delta_{ga}} \frac{1}{(\xi + i\sigma)(\xi - i\sigma)} \frac{1}{\Delta_{ga} - \xi} \frac{1}{\Delta_{ba} - \xi} \left(\xi - \Delta_{ga}\right)\\
			&= 2 \eta_{ijkl} i \sigma \left(\frac{1}{2 i \sigma} \frac{1}{-i \sigma - \Delta_{ga}} \frac{1}{\Delta_{ba} - i\sigma} \right) - 2\eta_{ijkl} i \sigma \left(\frac{1}{\sigma^2 + \Delta_{ga} ^2} \frac{1}{\Delta_{ga} + \Delta_{ba}} \right)\\
			&= -\frac{\eta_{ijkl}}{\Delta_{ga} + i \sigma} \left(\frac{1}{\Delta_{ba} - i \sigma} + \frac{2i\sigma}{(\Delta_{ga} - i \sigma)(\Delta_{ga} + \Delta_{ba})}\right)\\
		\end{split}
	\end{equation}
\end{widetext}
